On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}

### 2020-04-08 formal_glossary.tex
overall review (minor changes)

As mentioned before, the accessibility relation $\modalrelation$ is free from
any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  We
note that we have defined the Causality principle without considering it a
\emph{threat source} since, given that this is the first and outermost concept
in Figure~\ref{fig:security-safety}, we lack the concept of intent and
maliciousness (which are not inherent to the Causality Principle). 
In other words, the Causality principle or the Kripke structure
doesn't correlate cause-effect (or worlds) due to any other principle but
causality itself; therefore, in order to distinguish nominal or honest cause-effects
and non-nominal or dishonest ones we would need to add additional constraints, defining
what honesty is in a factious way.

@@

\Paragraph{Attacker as a Dis-honest Agent}
As we mentioned in Section~\ref{sec:problem_statement} when we postulated the
security hypothesis, Weaknesses and Vulnerability are the two core security
concepts of the current understanding of what security is. However, we see two
(opposed in terms of abstraction) possibility to better defined those concepts:
defining how an agent (Mankind or Nature) can maliciously use Weaknesses to
produce Vulnerability (and, in turn, Exploits), or how Errors can be defined
(since Error is a term from which Weakness are, in turn, defined).  The former,
in our current formalization, implies detailing and constraining the Causality
Principle by introducing constraints on how agents acts. For example,
introducing the concept of intentionality, motivation, or desire (see the BDI
framework defined in \autocite{BDI} for a general overview of those concepts).
The symbolic approach (also called Dolev-Yao model) in the formal verification
of security and communication protocols, can be used in this spirit. The
attacker can be seen as an additional agent (often identified with the network
where the messages are exchanged) that sees the network traffic and manipulates
it. The actions are generic (the attacker can combine two messages, split a
message into its components, or decrypt/encrypt if it has the proper security
key) but not something that a dishonest agent cannot do. The main difference is
that an honest agent follows the nominal specification of the protocol while
the attacker doesn't. The attacker non-deterministically opens as many new
sessions as it wants and the causality structure in which this scenario
(protocol + attacker) has been defined, is searched for a state that is
considered erroneous with respect to some security requirements decided a
priori. Therefore, as we mentioned in Section~\ref{sec:problem_statement}, the
formal verification of the security of protocols can be considered a scientific
security theory; the implementations resulting from the formalized (and
verified) design can be tested against the formalized security requirements.
So, it is not completely true that ``Thus, formal approaches offer no escape
from our basic problem; only by making unfalsifiable assumptions (about what an
attacker can do) will they allow derivation of a necessary condition'' (as
Herley describes in\autocite{Herley2009so}) rather, the
unfalsifiability-related issue seems to stem from how security conditions are
defined, i.e. how security requirements are defined. One can easily see that
this is not just an issue in the formal protocol verification community but the
very foundations of the concept of Weakness and Vulnerability are defined in a
way that makes it impossible (they are necessary conditions) to make a general
security statement on any system (even if we tested every single Weakness on
the CWE on our system it may be the case that a Weakness han yet to be
discovered). Therefore, we hypothesize that a security theory should define in
a complete way how errors (and then Weaknesses, Vulnerability, and Exploits)
may be introduced.

>>>>

As mentioned before, the accessibility relation $\modalrelation$ is free from
any axiomatic restriction (e.g. it's not reflexive nor anti-reflexive).  We
note that we have defined the Causality principle without considering it a
\emph{threat source} since, given that this is the first and outermost concept
in Figure~\ref{fig:safety-security}, we lack the concept of intent and
maliciousness (which are not inherent to the Causality Principle). 
In other words, the Causality principle or the Kripke structure
doesn't correlate cause-effect (or worlds) due to any other principle but
causality itself; therefore, in order to distinguish nominal or honest cause-effects
and non-nominal or dishonest ones we would need to add additional constraints, defining
what honesty is in a factious way.

@@

\Paragraph{Attacker as a Dis-honest Agent}
As we mentioned in Section~\ref{sec:problem_statement} when we postulated the
security hypothesis, Weaknesses and Vulnerability are the two core security
concepts of the current understanding of what security is. However, we see two
(opposed in terms of abstraction) possibility to better defined those concepts:
defining how an agent (Mankind or Nature) can maliciously use Weaknesses to
produce Vulnerability (and, in turn, Exploits), or how Errors can be defined
(since Error is a term from which Weakness are, in turn, defined).  The former,
in our current formalization, implies detailing and constraining the Causality
Principle by introducing constraints on how agents acts. For example,
introducing the concept of intentionality, motivation, or desire (see the BDI
framework defined in \autocite{BDI} for a general overview of those concepts).
The symbolic approach (also called Dolev-Yao model) in the formal verification
of security and communication protocols, can be used in this spirit. The
attacker can be seen as an additional agent (often identified with the network
where the messages are exchanged) that sees the network traffic and manipulates
it. The actions are generic (the attacker can combine two messages, split a
message into its components, or decrypt/encrypt if it has the proper security
key) but not something that a dishonest agent cannot do. The main difference is
that an honest agent follows the nominal specification of the protocol while
the attacker doesn't. The attacker non-deterministically opens as many new
sessions as it wants and the causality structure in which this scenario
(protocol + attacker) has been defined, is searched for a state that is
considered erroneous with respect to some security requirements decided a
priori. Therefore, as we mentioned in Section~\ref{sec:problem}, the
formal verification of the security of protocols can be considered a scientific
security theory; the implementations resulting from the formalized (and
verified) design can be tested against the formalized security requirements.
So, it is not completely true that ``Thus, formal approaches offer no escape
from our basic problem; only by making unfalsifiable assumptions (about what an
attacker can do) will they allow derivation of a necessary condition'' (as
Herley describes in\autocite{Herley2009so}) rather, the
unfalsifiability-related issue seems to stem from how security conditions are
defined, i.e. how security requirements are defined. One can easily see that
this is not just an issue in the formal protocol verification community but the
very foundations of the concept of Weakness and Vulnerability are defined in a
way that makes it impossible (they are necessary conditions) to make a general
security statement on any system (even if we tested every single Weakness on
the CWE on our system it may be the case that a Weakness han yet to be
discovered). Therefore, we hypothesize that a security theory should define in
a complete way how errors (and then Weaknesses, Vulnerability, and Exploits)
may be introduced.
